---
title: "Class 9, Group Proj: Virginia Census"
author: "Heather, Lucas, Zach, Alex, and Scott"
date: "11/2/2020"
output: 
  html_document:
    df_print: paged
    fig_height: 5
    fig_width: 8.5
    highlight: default
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: yes
    table: kable
---

DataViz: Map Building Demo
Heather Kitada Smalley
Motivating Example
In this demo we will be looking at factors for housing Fairfax County, VA.

# Polygon Maps
```{r}
library(tidyverse)

#install.packages("maps")
library(maps)

va_counties <- map_data("county", "virginia") %>% 
  select(lon = long, lat, group, id = subregion)

head(va_counties)
```

## Points
```{r}
ggplot(va_counties, aes(lon, lat)) + 
  geom_point(size = .25, show.legend = FALSE) +
  coord_quickmap()
```


## Polygons
```{r}
ggplot(va_counties, aes(lon, lat, group = group)) +
  geom_polygon(fill = "white", colour = "grey50") + 
  coord_quickmap()
```

# Simple Features Maps
```{r}
#install.packages("sf")
library(sf)
```

```{r}
#install.packages("tigris")
library(tigris)
options(tigris_use_cache = TRUE)

# set year for boundaries 
this.year = 2017 #the last census

va_tracts <- tracts(state = 'VA', county = 'Fairfax County', 
                    cb = T, year = this.year)

# head(va_tracts)
```

## BASE PLOT
```{r}
plot(va_tracts)
```

## GGPLOT
```{r}
ggplot(va_tracts) + 
  geom_sf() + 
  coord_sf()
```

# Getting Data from the US Census Bureau
First you will need to get a API key: https://api.census.gov/data/key_signup.html

Once you have an API key load it into your R environment so that you can access the ACS data.
```{r,  include=FALSE}
#install.packages("tidycensus")
library(tidycensus)
## Warning: package 'tidycensus' was built under R version 3.6.2
# YOUR CODE SHOULD LOOK LIKE THIS
census_api_key("") # <- INSERT YOUR OWN API KEY
```


To install your API key for use in future sessions, run this function with `install = TRUE`.
Many many variables are included in the ACS. The ACS has 1 and 5 year estimates. Use the following code to see what variables are available:
```{r}
# Set a year of interest
this.year = 2010

# This looks at the 5 year estimates
# You can also do "acs1"
vars <- load_variables(year = this.year,
                      dataset = "acs5",
                      cache = TRUE)

# There are 25070 possible variables 
dim(vars)
```

Explore several possible explantory variables from the American Community Survey (ACS) including:

B02001_001: Total
B03002_003: White alone (Not Hispanic or Latino)
B03002_004 Black or African American alone (Not Hispanic or Latino)
B03002_012: Hispanic or Latino
B03002_005: Native American alone (Not Hispanic or Latino)
B03002_006: Asian alone (Not Hispanic or Latino)
B03002_007: Native Hawaiian or Pacific Islander alone (Not Hispanic or Latino)
B03002_009: Multiple Races (Not Hispanic or Latino)
B03002_008: Other (Not Hispanic or Latino)
B25064_001 MEDIAN GROSS RENT
B25071_001: Rent Burden (MEDIAN GROSS RENT AS A PERCENTAGE OF HOUSEHOLD INCOME)
B19013_001: MEDIAN HOUSEHOLD INCOME IN PAST 12 MONTHS
B01002_001: Median age
B25115_016: Renter Occupied - family
B25115_027: Renter Occupied - nonfamily

## Getting data for one variable
MEDIAN HOME VALUE
```{r}
vaMedv <- get_acs(geography = "tract", year=this.year,
               state = "VA", county = "Fairfax County",
               variables = "B25077_001E")
head(vaMedv)
```

Notice that we’re going to need to do a little data wrangling so that we have a tidydata format to spread the column named ‘variable’.

## Getting data for multiple variables 
Get census tract level estimates for Fairfax County:
```{r}
## Names for variable types
# Gives five year estimates
vaHouse <- get_acs(geography = "tract", year=this.year,
                  state = "VA", county = "Fairfax County", geometry = TRUE,
                  variables = c(popululation = "B02001_001",
                                median.gross.rent = "B25064_001",
                                median.household.income = "B19013_001",
                                rent.burden = "B25071_001"))
## Getting data from the 2006-2010 5-year ACS
head(vaHouse)
```

```{r}
vaTidy<-as.data.frame(vaHouse)[,c(1,3:4)]%>%
  spread(variable, estimate)

# head(vaTidy)
```

# Geojoins and TMAPS
Suppose that we want to study median house value (which is B25077_001E).

```{r}
vaMedv <- get_acs(geography = "tract", year=this.year,
              state = "VA", county = "Fairfax County",
              variables = "B25077_001E")%>%
  mutate(AFFGEOID=paste0("1400000US", GEOID))
```

Getting data from the 2006-2010 5-year ACS
In order to combine data from the US Census with our spatial data from trigis we will need to use a geo_join.

```{r}
joinVA<-geo_join(va_tracts, vaMedv, 
                 by_sp="AFFGEOID", by_df="AFFGEOID")
```


Then we will use the tmap package to plot the data with a green color gradient (with 7 levels).
```{r}
#install.packages("tmap")
library(tmap)
## Warning: package 'tmap' was built under R version 3.6.2
## USE TMAP PACKAGE
tm_shape(joinVA, projection = 26916)+
  tm_fill("estimate", style = "quantile", n=7, palette = "Greens")+
  tm_legend(bg.color="white", bg.alpha=0.6)
```




# Interactive Maps with Leaflet
## Pop-ups
```{r}
# MEDIAN HOME VALUE
vaMedvG <- get_acs(geography = "tract", year=this.year,
               state = "VA", county = "Fairfax County",
               variables = "B25077_001E", 
               geometry = TRUE)
```

```{r}
#install.packages("leaflet")
library(leaflet)
```

```{r}
pal<-colorNumeric("Greens", domain=0:ceiling(max(vaMedvG$estimate, na.rm=TRUE)))

popup<-paste("Tract: ", as.character(substring(vaMedvG$GEOID, 6, 11)), "<br>",
             "Median Home Value: ", as.character(vaMedvG$estimate))

leaflet()%>%
  addProviderTiles("CartoDB.Positron")%>%
  addPolygons(data=vaMedvG,
              fillColor= ~pal(vaMedvG$estimate),
              fillOpacity = .7,
              weight =.5,
              smoothFactor = 0.2,
              popup = popup)

```

## Quantile Color Palette
```{r}
qpal<-colorQuantile("viridis", domain=vaMedvG$estimate,
                       n=5,na.color="#FFFFFF")

leaflet()%>%
  addProviderTiles("CartoDB.Positron")%>%
  addPolygons(data=vaMedvG,
              fillColor= ~qpal(vaMedvG$estimate),
              fillOpacity = 0.7,
                  color="grey",
                  opacity=.5,
                  weight = 0.4,
                  smoothFactor = 0.2,
                  popup = popup)%>%
      addLegend("bottomright", pal=qpal, values=vaMedvG$estimate,
                opacity = .7,
                title="Percentiles")
```
