---
title: 'DATA 502: Midterm 2 Part 3'
author: "Heather Rugnetta"
date: "11/22/2020"
output:
  html_document:
    df_print: paged
    fig_height: 5
    fig_width: 8.5
    highlight: default
    number_sections: yes
    theme: journal
    toc: yes
    toc_depth: 2
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  word_document:
    toc: yes
    toc_depth: '2'
---

# Option 1: Create
## Step 1: Use the variables from the dataset to ask a question about Thanksgiving
### Who eats coconut pie for Thanksgiving?
### Who eats (fruit) salad for Thanksgiving?

# Setup
## Load the Libraries
```{r  setup, include=FALSE}
library(tidyverse)
library(ggplot2)
library(dplyr)
library(usmap)
library(maps)
library(xkcd)
library(sf)
library(maptools)
library(rgeos)
library (plotly)
```

## Import data
```{r import}
thanks <-read.csv( "https://raw.githubusercontent.com/kitadasmalley/FA2020_DataViz/main/data/useThanks.csv",
header=TRUE, na.strings = c("","NA"))
head(thanks)
```
```{r}
thanks <- thanks[!is.na(thanks$DivName), ]
```

## Finagle data (in a good way)
### Count Coconut Fans
```{r coconut count}
temp1 <- thanks%>% 
  group_by(DivName) %>% 
  summarize(Coconut = sum(coconut.pie01))

```
### Count the Fruit Salad peeps
```{r fruit salad count}
temp2 <- thanks%>% 
  group_by(DivName) %>% 
  summarize(FruitSalad = sum(fruit.salad01))

```
```{r who}
temp3 <- thanks%>% 
  group_by(DivName) %>% 
  summarize(Respondants = sum(celebrate01))

```

### Join the results of the Coconut and Fruit Salad fans
```{r join fruity folks}
temp4 <- inner_join(temp1, temp2)
temp5 <- inner_join(temp4, temp3)
temp5 <- temp5 %>% 
   mutate(DivName=paste(DivName, " Division", sep=""))
temp5
```
```{r}
temp6 <- temp5 %>% 
  group_by(DivName) %>% 
  summarize(propCoconut= Coconut/Respondants,
         propFruit=FruitSalad/Respondants)
temp6
```

## Get the US Map and States data
```{r Map data}
states <- usmap::us_map()
colnames(states)[9] <- "Name"
```

```{r}
fips<-read.csv("https://raw.githubusercontent.com/kitadasmalley/FA2020_DataViz/main/data/stateFIPS.csv", 
               header=TRUE)
colnames(fips)[1] <- "Name"
#colnames(fips)[2] <- "state"
fips
```
```{r}
states <- usmap::us_map()
colnames(states)[9] <- "Name"
```

```{r}
stateFips <- states %>% select(fips, Name)
stateFips <-stateFips %>% unique()

```
```{r}
foodStates<-stateFips %>%
  left_join(fips)

```

## Join the States and Food data by Division Name
```{r join map and food data}
fruitStates<- foodStates %>% 
  left_join(temp6) 


```


# Plot
## Step 2: Make your first attempt at creating a graphic
Test out the US Map
```{r b&w of states}
plot_usmap()
```

## Initial look at Data
```{r}
plot_usmap(data= fruitStates, values ="propCoconut")
```


## Polishing

```{r polish coconut}
plot_usmap(data = fruitStates, values = "propCoconut", color = "black", labels=F) + 
  scale_fill_continuous( low = "white", high = "blueviolet", name = "response") + 
  labs(title = "Who Eats Coconut Pie for Thanksgiving? by State")
```
```{r}
plot_usmap(data = fruitStates, values = "propFruit", color = "black", labels=F) + 
  scale_fill_continuous( low = "white", high = "brown4", name = "Percentage", label = scales::percent) + 
  labs(title = "Who Eats Fruit Salad for Thanksgiving? by US Region", subtitle = "Percentage of 1058 Respondants in 2015")
```

```{r}
plot_usmap(data = fruitStates, values = "propFruit", color = "black", labels=F) + 
  scale_fill_continuous( low = "white", high = "darkred", name = "% of Respondants", label = scales::label_percent(2))+
  theme (legend.position = "right")+
  labs(title = "Who Eats Fruit Salad for Thanksgiving? by US Region", subtitle = "Percentage of 1058 Respondants in 2015")

```

## Final Products
```{r}
plot_usmap(data = fruitStates, values = "propCoconut", color = "black", labels=F) + 
  scale_fill_continuous( low = "black", high = "cyan", name = "% of Respondants", label = scales::label_percent(2))+
  theme (legend.position = "right")+
  labs(title = "Who Eats Coconut Pie for Thanksgiving? ", subtitle = "Percentage of US Region Respondants*", caption= "*Survey of 1,058 Respondants in 2015")+
  theme(panel.background = element_rect(color = "black", fill = "lightblue"))

```
```{r}
f <- plot_usmap(data = fruitStates, values = "propFruit", color = "black", labels=F) + 
  scale_fill_continuous( low = "white", high = "darkred", name = "% of Respondants", label = scales::label_percent(2))+
  theme (legend.position = "right")+
  labs(title = "Who Eats Fruit Salad for Thanksgiving? by US Region",subtitle = "Percentage of US Region Respondants*", caption= "*Survey of 1,058 Respondants in 2015")+
  theme(panel.background = element_rect(color = "black", fill = "darkgoldenrod1"))
f <- ggplotly(f)
f
```

### Other things I wanted to do, but ran out of time
#### Create regional outlines 
#### Incorporate Leaflet and hovering
#### Compare with other obscure traditions in a "pie chart"

